---
title: "OMEGA"
output: html_document
---

```{r}
require(ggplot2)
require(plyr)
require(lubridate)
require(dplyr)
require(sp)
require(maptools)
require(ggmap)
require(ggplot2)
require(reshape2)
require(rgdal)
require(raster)
require(rgeos)
require(PBSmapping)
require(gdistance)
require(grid)
require(gridExtra)
require(scales)
```

### read in and tidy up
```{r}
setwd("C:/Users/keyserf/Documents/Data/Modelling/")
broodfitness <- read.csv("./OMEGA results_2016-12-14_AtlSalmon1_brood year fitness.csv", header=FALSE)
progescape <- read.csv("./OMEGA results_2016-12-14_AtlSalmon1_escapees.csv", header=FALSE)
runsumm <- read.csv("./OMEGA results_2016-12-14_AtlSalmon1_Summary.csv", header=FALSE)
allabund <- read.csv("./OMEGA results_2016-12-14_AtlSalmon1_all abund.csv", header=FALSE)
wildabund <- read.csv("./OMEGA results_2016-12-14_AtlSalmon1_wild abund.csv", header=FALSE)

runsumm[,1:10] <- apply(runsumm[,1:10], 2, function(x) as.character(x))
runsumm[,1:10] <- apply(runsumm[,1:10], 2, function(x) gsub(",", "", x))
cols <- runsumm[,1]
cols <- gsub(" ", ".", cols)

runsumm <- data.frame(cbind(c(runsumm[1, 2:10]), 
                      c(runsumm[2, 2:10]),
                      c(runsumm[3, 2:10]),
                      c(runsumm[4, 2:10]),
                      c(runsumm[5, 2:10]),
                      c(runsumm[6, 2:10]),
                      c(runsumm[7, 2:10])))

colnames(runsumm) <- cols

runsumm[,1:7] <- apply(runsumm[,1:7], 2, function(x) as.numeric(x))

ggplot() + geom_point(data=runsumm, aes(Age, Productivity.to.end.of.year), shape=1) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  xlab("Age") + 
  ylab("Productivity to end of year") +
  xlim(0,10) 

ggplot() + geom_point(data=runsumm, aes(Age, Fishing.Mortality), shape=2) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  xlab("Age") + 
  ylab("Fishing mortality") +
  xlim(0,10) 

par(mar = c(5,5,2,5))
plot(runsumm$Age, runsumm$Productivity.to.end.of.year, xlim=c(0,10), xlab=NA, ylab=NA, cex=1.2)
par(new = T)
plot(runsumm$Age, runsumm$Fishing.Mortality, xlim=c(0,10), pch=16, axes=F, xlab=NA, ylab=NA, cex=1.2)
axis(side = 4)
mtext(side = 4, line = 3, 'Fishing mortality')
mtext(side = 2, line = 3, 'Productivity to end of year')
mtext(side = 1, line = 3, 'Age')
legend("bottomright",
       legend=c("Productivity to end of year", "Fishing mortality"),
       pch=c(1, 16))
title("Atlantic salmon run 1")



allabund <- allabund[2:101,]
colnames(allabund) <- c("Year", "Biomass_Fem_Spawn_ALL", 0:10)

allabund[,1:13] <- apply(allabund[,1:13], 2, function(x) gsub(",", "", x))
allabund[,1] <- gsub("Year ", "", allabund[,1])
allabund[,1:13] <- apply(allabund[,1:13], 2, function(x) as.numeric(x))
head(allabund)

allabund <- melt(allabund, id.vars=c("Year", "Biomass_Fem_Spawn_ALL"))
allabund$variable <- as.numeric(as.character(allabund$variable))


ggplot() + geom_point(data=allabund[allabund$variable>0,], aes(Year, value, colour=as.factor(variable))) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  xlab("Year") +
  ylab("log(Total Abundance)") +
  scale_colour_discrete(name="Age") +
  ggtitle("Atlantic salmon run 1\nTotal Abundance by age")

ggplot() + geom_point(data=allabund, aes(Year, Biomass_Fem_Spawn_ALL)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  ylab("Total Female Spawner Biomass (mt)") +
  ggtitle("Atlantic salmon run 1")


wildabund <- wildabund[2:101,1:13]
colnames(wildabund) <- c("Year", "Biomass_Fem_Spawn_W", 0:10)

wildabund[,1:13] <- apply(wildabund[,1:13], 2, function(x) gsub(",", "", x))
wildabund[,1] <- gsub("Year ", "", wildabund[,1])
wildabund[,1:13] <- apply(wildabund[,1:13], 2, function(x) as.numeric(x))
head(wildabund)

wildabund <- melt(wildabund, id.vars=c("Year", "Biomass_Fem_Spawn_W"))
wildabund$variable <- as.numeric(as.character(wildabund$variable))
# 
# wildabund$size[wildabund$variable <5] <- "1-5"
# wildabund$size[wildabund$variable > 4 & wildabund$variable < 10] <- "5-9"
# wildabund$size[wildabund$variable > 9 & wildabund$variable < 15] <- "10-14"
# wildabund$size[wildabund$variable > 14 & wildabund$variable < 20] <- "15-19"
# wildabund$size[wildabund$variable > 19 & wildabund$variable < 25] <- "20-24"
# wildabund$size[wildabund$variable > 24 & wildabund$variable < 30] <- "25-29"
# wildabund$size[wildabund$variable > 29 & wildabund$variable < 40] <- "30-39"
# wildabund$size[wildabund$variable > 39 & wildabund$variable < 50] <- "40-49"
# wildabund$size[wildabund$variable > 49 & wildabund$variable < 60] <- "50-59"
# wildabund$size[wildabund$variable > 59 & wildabund$variable < 70] <- "60-69"
# 
# wildabund$size <- factor(wildabund$size, levels=c("1-5", "5-9", "10-14", "15-19", "20-24", "25-29", "30-39", "40-49", "50-59", "60-69"))

ggplot() + geom_point(data=wildabund[wildabund$variable>0,], aes(Year, log(value), colour=as.factor(variable))) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  xlab("Year") +
  ylab("log(Wild Abundance)") +
  scale_colour_discrete(name="Age") +
  ggtitle("Atlantic salmon run 1\nWild Abundance by age bin")

ggplot() + geom_point(data=wildabund, aes(Year, Biomass_Fem_Spawn_W)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  ylab("Wild Female Spawner Biomass (mt)") +
  ggtitle("Sablefish example")


names(wildabund) <- c("Year", "Wild_spawn_biomass", "Age", "Wild_abund")
names(allabund) <- c("Year", "Total_spawn_biomass", "Age", "Total_abund")

abund <- join(wildabund, allabund, type="left")

abund$Escape_spawn_biomass <- abund$Total_spawn_biomass - abund$Wild_spawn_biomass
abund$Escape_abund <- abund$Total_abund - abund$Wild_abund

biomass <- melt(abund[,c(1:3, 6)], id.vars=c("Year", "Age"))

abund2 <- melt(abund[,c(1,3, 4, 7, 9)], id.vars=c("Year", "Age"))

levels(biomass$variable) <- c("Wild", "Total") 

ggplot() + geom_point(data=biomass, aes(Year, log(value), shape=variable)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  scale_shape_manual(values=c(1,16), name=NULL) +
  ylab("Female Spawner biomass (mt)") +
  ggtitle("Sablefish example")

require(scales)

ggplot() + geom_point(data=abund2[abund2$variable %in% c("Escape_abund"),], aes(Year, value)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  ylab("Escapee abundance (mt)\n") + 
  xlab("\nYear") +
  scale_y_continuous(labels = comma) +
  ggtitle("Sablefish example")
  


cols <- progescape[1,]
progescape <- progescape[2:101,]
colnames(progescape) <- c("Year", "Inventory", "Escapees", "Weight_bin1", "Weight_bin1.5", "Weight_bin2.0")

progescape <- progescape[,1:6]

progescape[,1:6] <- apply(progescape[,1:6], 2, function(x) gsub(",", "", x))
progescape[,1] <- gsub("Year ", "", progescape[,1])
progescape[,1:6] <- apply(progescape[,1:6], 2, function(x) as.numeric(x))
head(progescape)

progescape <- melt(progescape, id.vars=c("Year", "Inventory", "Escapees"))

ggplot() + geom_point(data=progescape[progescape$variable == "Weight_bin1",], aes(Year, Escapees)) +
  theme_classic() + 
  theme(panel.background=element_rect(colour="black")) +
  scale_y_continuous(labels=comma) +
  ylab("Number of escapees\n") +
  xlab("\nYear") +
  ggtitle("Atlantic salmon run 1")



cols <- broodfitness[1,]
broodfitness <- broodfitness[2:101,]
colnames(broodfitness) <- c("Year", "LifeCycle", "NaturalTraitValue", "EscapeeTraitValue", "pHOS")

broodfitness <- broodfitness[,1:5]

broodfitness[,1] <- gsub("Year ", "", broodfitness[,1])
broodfitness[,1:5] <- apply(broodfitness[,1:5], 2, function(x) as.numeric(x))
head(broodfitness)

ggplot() + geom_point(data=broodfitness, aes(Year, LifeCycle)) +
  theme_classic() + 
  theme(panel.background = element_rect(colour="black"))

broodfitness <- melt(broodfitness, id.vars=c("Year", "LifeCycle", "pHOS"))

ggplot() + geom_point(data=broodfitness[broodfitness$variable=="NaturalTraitValue",], aes(Year, value)) +
  theme_classic() + 
  theme(panel.background = element_rect(colour="black")) +
  ylim(80,100) +
  ylab("Natural Trait Value\n") +
  xlab("\nYear") +
  annotate(geom="text", x=85, y=99.75, label="Brood year fitness", size=5)


ggplot() + geom_point(data=broodfitness[broodfitness$variable=="NaturalTraitValue",], aes(Year, pHOS)) +
  theme_classic() + 
  theme(panel.background = element_rect(colour="black")) +
  ylab("Percentage of hatchery-origin spawners\n") +
  xlab("\nYear") +
  ggtitle("Sablefish example\nBrood year fitness") +
  scale_y_continuous(label=percent)
```



### calculating parameters
```{r}
annual <- read.csv("C:/Users/keyserf/Documents/Data/Inventory data/NL annual inventory and coords_summary_2016-11-29.csv")

plot(annual$Long, annual$Lat)
#click(n=4)
points(CRAQ$Long, CRAQ$Lat, col="red", add=TRUE)

CRAQ <- subset(annual, Long > -56.25 & Long < -55.6 & Lat > 47.584 & Lat < 47.94)

CRAQ2013 <- subset(CRAQ, ReportYear=="2013")

CRAQ2013$tonnes <- CRAQ2013$totalfish * 0.002
mean(CRAQ2013$tonnes)
sum(CRAQ2013$tonnes)
sum(CRAQ$totalfish)

allNL <- read.csv("C:/Users/keyserf/Documents/Data/Inventory data/NL inventory master2_fixed.csv", as.is=TRUE)

NL2013 <- subset(allNL, Reporting.Year=="2013")
str(NL2013)

NL2013 <- ddply(.data=NL2013, .(Site.AQ),
                summarize,
                totmorts = sum(as.numeric(No..Fish.Morts), na.rm=TRUE),
                totstart = sum(as.numeric(Starting.No..Fish), na.rm=TRUE),
                totadd = sum(as.numeric(No..Fish.Introduced), na.rm=TRUE))
NL2013$surv <- NL2013$totmorts/(NL2013$totstart+NL2013$totadd)

NL2013$surv2 <- 1-NL2013$surv
NL2013 <- NL2013[c(1:18, 21:33),]
mean(NL2013$surv2, na.rm=TRUE)



# L-W relationship
#smolt mean weight = 31.8 g, smolt mean size = 14.9 cm
#small mean weight = 1468 g, small mean size = 51.6 cm
#large mean weight = 3016 g, large mean size = 68.9 cm

length <- c(14.9, 51.6, 68.9)
weight <- c(31.8, 1468, 3016)/1000
LW <- data.frame(cbind(length, weight))
lm(data=LW, length~weight)

cages2013 <- subset(allNL, Reporting.Year=="2013")
length(unique(cages2013$Cage.No.))
