---
title: "OMEGA"
output: html_document
---

```{r}
require(ggplot2)
require(plyr)
require(lubridate)
require(dplyr)
require(sp)
require(maptools)
require(ggmap)
require(ggplot2)
require(reshape2)
require(rgdal)
require(raster)
require(rgeos)
require(PBSmapping)
require(gdistance)
require(grid)
require(gridExtra)
require(scales)
```

### read in and tidy up
```{r}
setwd("C:/Users/keyserf/Documents/Data/")
broodfitness <- read.csv("./OMEGA results_2016-11-04_Sablefish1_Brood year fitness.csv", header=FALSE)
progescape <- read.csv("./OMEGA results_2016-11-04_Sablefish1_Prog escapees by bin.csv", header=FALSE)
runsumm <- read.csv("./OMEGA results_2016-11-04_Sablefish1_Summary.csv", header=FALSE)
fembiomass <- read.csv("./OMEGA results_2016-11-04_Sablefish1_F Biomass wild and escapee.csv", header=FALSE)
wildabund <- read.csv("./OMEGA results_2016-11-04_Sablefish1_Abundance_wild.csv", header=FALSE)

runsumm[,1:67] <- apply(runsumm[,1:67], 2, function(x) as.character(x))
runsumm[,1:67] <- apply(runsumm[,1:67], 2, function(x) gsub(",", "", x))
cols <- runsumm[,1]
cols <- gsub(" ", ".", cols)

runsumm <- data.frame(cbind(c(runsumm[1, 2:67]), 
                      c(runsumm[2, 2:67]),
                      c(runsumm[3, 2:67]),
                      c(runsumm[4, 2:67]),
                      c(runsumm[5, 2:67]),
                      c(runsumm[6, 2:67]),
                      c(runsumm[7, 2:67])))

colnames(runsumm) <- cols

runsumm[,1:7] <- apply(runsumm[,1:7], 2, function(x) as.numeric(x))

ggplot() + geom_point(data=runsumm, aes(Age, Productivity.to.end.of.year), shape=1) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  xlab("Age") + 
  ylab("Productivity to end of year") +
  xlim(0,20) 

ggplot() + geom_point(data=runsumm, aes(Age, Fishing.Mortality), shape=2) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  xlab("Age") + 
  ylab("Fishing mortality") +
  xlim(0,20) 

par(mar = c(5,5,2,5))
plot(runsumm$Age, runsumm$Productivity.to.end.of.year, xlim=c(0,20), xlab=NA, ylab=NA, cex=1.2)
par(new = T)
plot(runsumm$Age, runsumm$Fishing.Mortality, xlim=c(0,20), pch=16, axes=F, xlab=NA, ylab=NA, cex=1.2)
axis(side = 4)
mtext(side = 4, line = 3, 'Fishing mortality')
mtext(side = 2, line = 3, 'Productivity to end of year')
mtext(side = 1, line = 3, 'Age')
legend("bottomright",
       legend=c("Productivity to end of year", "Fishing mortality"),
       pch=c(1, 16))
title("Sablefish example")



fembiomass <- fembiomass[2:101,]
colnames(fembiomass) <- c("Year", "Biomass_Fem_Spawn_ALL", 0:65)

fembiomass[,1:68] <- apply(fembiomass[,1:68], 2, function(x) gsub(",", "", x))
fembiomass[,1] <- gsub("Year ", "", fembiomass[,1])
fembiomass[,1:68] <- apply(fembiomass[,1:68], 2, function(x) as.numeric(x))
head(fembiomass)

fembiomass <- melt(fembiomass, id.vars=c("Year", "Biomass_Fem_Spawn_ALL"))
fembiomass$variable <- as.numeric(as.character(fembiomass$variable))

fembiomass$size[fembiomass$variable <5] <- "1-5"
fembiomass$size[fembiomass$variable > 4 & fembiomass$variable < 10] <- "5-9"
fembiomass$size[fembiomass$variable > 9 & fembiomass$variable < 15] <- "10-14"
fembiomass$size[fembiomass$variable > 14 & fembiomass$variable < 20] <- "15-19"
fembiomass$size[fembiomass$variable > 19 & fembiomass$variable < 25] <- "20-24"
fembiomass$size[fembiomass$variable > 24 & fembiomass$variable < 30] <- "25-29"
fembiomass$size[fembiomass$variable > 29 & fembiomass$variable < 40] <- "30-39"
fembiomass$size[fembiomass$variable > 39 & fembiomass$variable < 50] <- "40-49"
fembiomass$size[fembiomass$variable > 49 & fembiomass$variable < 60] <- "50-59"
fembiomass$size[fembiomass$variable > 59 & fembiomass$variable < 70] <- "60-69"

fembiomass$size <- factor(fembiomass$size, levels=c("1-5", "5-9", "10-14", "15-19", "20-24", "25-29", "30-39", "40-49", "50-59", "60-69"))

ggplot() + geom_point(data=fembiomass[fembiomass$variable>0,], aes(Year, log(value), colour=variable)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  facet_wrap(~size, nrow=2) +
  xlab("Year") +
  ylab("log(Total Abundance)") +
  scale_colour_continuous(name="Age") +
  ggtitle("Sablefish example\nTotal Abundance by age bin")

ggplot() + geom_point(data=fembiomass, aes(Year, Biomass_Fem_Spawn_W)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  ylab("Total Female Spawner Biomass (mt)") +
  ggtitle("Sablefish example")


wildabund <- wildabund[3:102,1:68]
colnames(wildabund) <- c("Year", "Biomass_Fem_Spawn_W", 0:65)

wildabund[,1:68] <- apply(wildabund[,1:68], 2, function(x) gsub(",", "", x))
wildabund[,1] <- gsub("Year ", "", wildabund[,1])
wildabund[,1:68] <- apply(wildabund[,1:68], 2, function(x) as.numeric(x))
head(wildabund)

wildabund <- melt(wildabund, id.vars=c("Year", "Biomass_Fem_Spawn_W"))
wildabund$variable <- as.numeric(as.character(wildabund$variable))

wildabund$size[wildabund$variable <5] <- "1-5"
wildabund$size[wildabund$variable > 4 & wildabund$variable < 10] <- "5-9"
wildabund$size[wildabund$variable > 9 & wildabund$variable < 15] <- "10-14"
wildabund$size[wildabund$variable > 14 & wildabund$variable < 20] <- "15-19"
wildabund$size[wildabund$variable > 19 & wildabund$variable < 25] <- "20-24"
wildabund$size[wildabund$variable > 24 & wildabund$variable < 30] <- "25-29"
wildabund$size[wildabund$variable > 29 & wildabund$variable < 40] <- "30-39"
wildabund$size[wildabund$variable > 39 & wildabund$variable < 50] <- "40-49"
wildabund$size[wildabund$variable > 49 & wildabund$variable < 60] <- "50-59"
wildabund$size[wildabund$variable > 59 & wildabund$variable < 70] <- "60-69"

wildabund$size <- factor(wildabund$size, levels=c("1-5", "5-9", "10-14", "15-19", "20-24", "25-29", "30-39", "40-49", "50-59", "60-69"))

ggplot() + geom_point(data=wildabund[wildabund$variable>0,], aes(Year, log(value), colour=variable)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  facet_wrap(~size, nrow=2) +
  xlab("Year") +
  ylab("log(Wild Abundance)") +
  scale_colour_continuous(name="Age") +
  ggtitle("Sablefish example\nWild Abundance by age bin")

ggplot() + geom_point(data=wildabund, aes(Year, Biomass_Fem_Spawn_W)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  ylab("Wild Female Spawner Biomass (mt)") +
  ggtitle("Sablefish example")


names(wildabund) <- c("Year", "Wild_spawn_biomass", "Age", "Wild_abund", "Size")
names(fembiomass) <- c("Year", "Total_spawn_biomass", "Age", "Total_abund", "Size")

abund <- join(wildabund, fembiomass, type="left")

abund$Escape_spawn_biomass <- abund$Total_spawn_biomass - abund$Wild_spawn_biomass
abund$Escape_abund <- abund$Total_abund - abund$Wild_abund

biomass <- melt(abund[,c(1:3, 6)], id.vars=c("Year", "Age"))

abund2 <- melt(abund[,c(1,3, 4, 7, 9)], id.vars=c("Year", "Age"))

levels(biomass$variable) <- c("Wild", "Total") 

ggplot() + geom_point(data=biomass, aes(Year, value, shape=variable)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  scale_shape_manual(values=c(1,16), name=NULL) +
  ylab("Female Spawner biomass (mt)") +
  ggtitle("Sablefish example")

ggplot() + geom_point(data=abund2[abund2$variable %in% c("Escape_abund"),], aes(Year, value)) +
  theme_classic() +
  theme(panel.background=element_rect(colour="black")) +
  ylab("Escapee abundance (mt)\n") + 
  xlab("\nYear") +
  scale_y_continuous(labels = comma) +
  ggtitle("Sablefish example")
  


cols <- progescape[1,]
progescape <- progescape[2:101,]
colnames(progescape) <- c("Year", "Inventory", "Escapees", "Weight_bin.015", "Weight_bin.2", "Weight_bin.4", "Weight_bin.6", "Weight_bin.8", "Weight_bin1.")

progescape[,1:9] <- apply(progescape[,1:9], 2, function(x) gsub(",", "", x))
progescape[,1] <- gsub("Year ", "", progescape[,1])
progescape[,1:9] <- apply(progescape[,1:9], 2, function(x) as.numeric(x))
head(progescape)

progescape <- melt(progescape, id.vars=c("Year", "Inventory", "Escapees"))

ggplot() + geom_point(data=progescape[progescape$variable == "Weight_bin1.",], aes(Year, Escapees)) +
  theme_classic() + 
  theme(panel.background=element_rect(colour="black")) +
  scale_y_continuous(labels=comma) +
  ylab("Number of escapees\n") +
  xlab("\nYear") +
  ggtitle("Sablefish example")



cols <- broodfitness[1,]
broodfitness <- broodfitness[2:101,]
colnames(broodfitness) <- c("Year", "LifeCycle", "NaturalTraitValue", "EscapeeTraitValue", "pHOS")

broodfitness[,1] <- gsub("Year ", "", broodfitness[,1])
broodfitness[,1:5] <- apply(broodfitness[,1:5], 2, function(x) as.numeric(x))
head(broodfitness)

ggplot() + geom_point(data=broodfitness, aes(Year, LifeCycle)) +
  theme_classic() + 
  theme(panel.background = element_rect(colour="black"))

broodfitness <- melt(broodfitness, id.vars=c("Year", "LifeCycle", "pHOS"))

ggplot() + geom_point(data=broodfitness, aes(Year, value, shape=variable)) +
  theme_classic() + 
  theme(panel.background = element_rect(colour="black")) +
  scale_shape_manual(values=c(1,16), name=NULL, labels=c("Natural", "Escapee")) +
  ylim(70,100) +
  ylab("Trait Value\n") +
  xlab("\nYear") +
  ggtitle("Sablefish example\nBrood year fitness")


ggplot() + geom_point(data=broodfitness[broodfitness$variable=="NaturalTraitValue",], aes(Year, pHOS)) +
  theme_classic() + 
  theme(panel.background = element_rect(colour="black")) +
  ylab("Percentage of hatchery-origin spawners\n") +
  xlab("\nYear") +
  ggtitle("Sablefish example\nBrood year fitness") +
  scale_y_continuous(label=percent)

